---
title: "Lab5"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output: html_document
---

```{r packages, include = FALSE}
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) 
library(broom) 
library(textdata)
library(ggplot2)
library(dplyr)
library(LexisNexisTools)
```


# Assignment #5
## Train Your Own Embeddings

1.  Using the data from your Nexis Uni query from Week 2, create a set of word embeddings. To do this, you'll essentially need to recreate the steps in today's example down through the chunk named "pmi".

```{r}
text <- readRDS(here::here("lab2_text")) |> 
  janitor::clean_names()
```

#### a) Creating a data frame with all the words of interest, and calculate the unigram probabilities. 
```{r}
# Unigram is a single word level of a text
unigram_probs <- text |> 
  unnest_tokens(word, text) |> 
  anti_join(stop_words, by = "word") |> # Delete words that are not useful
  count(word, sort = T) |> 
  mutate(p = n/sum(n)) #Probability of each word

```


2.  Think of 3-5 key words in your data set. Calculate and plot the 10 most semantically similar words for each of them. Identify and interpret any interesting or surprising results.

3.  Assemble 3 word math equations that you think could be useful or interesting for exploring the meaning of key words or ideas in your data set.

#### Pretrained Embeddings

4.  Following the example in the SMLTR text (section 5.4), create a set of 100-dimensional GloVe word embeddings. These embeddings were trained by researchers at Stanford on 6 billion tokens from Wikipedia entries.

5.  Test them out with the cannonical word math equation on the GloVe embeddings: "berlin" - "germany" + "france" = ?

6.  Recreate parts 2 and 3 above using the the GloVe embeddings in place of the ones you trained. How do they compare? What are the implications for applications of these embeddings?